Length Gauge Theory of Surface Second Harmonic Generation

### 2014-11-17 intro.tex
removed commas per abhishek

Numerous mathematical frameworks have been developed to study the 
nonlinear optical response of materials; 
in the context of this paper, we note two in particular. 
The first, called the ``velocity gauge'' or 
``$\mathbf{v}\cdot\mathbf{A}$ gauge,'' features in early 
research in this field. 
Butcher and McLean\cite{butcherPOPS63}, and 
Aspnes\cite{aspnesPRB72} 
laid the initial groundwork by deriving explicit expressions for the 
nonlinear susceptibility tensor. However, these either diverged at 
low frequencies or were limited to zincblende crystal structures. 
A few decades later, 
Ghahramani \emph{et al}.\cite{ghahramaniPRB91} 
realized a full-band-structure calculation for Si/Ge superlattices and 
found significant surface SHG while providing proof that the expressions 
do not diverge at low frequencies for clean, cold semiconductors. 
Sipe \emph{et al}.\cite{sipePRB93}
separated the interband and intraband one-electron contributions 
of the susceptibility tensor and further verified the non-divergent behavior. 
Levine\cite{levinePRB94} 
introduced the scissors operator within the Local Density Approximation
(LDA) and found good agreement with experimental results. 
The second, called the ``length gauge'' or 
``$\mathbf{r}\cdot\mathbf{E}$ gauge'' has various advantages over 
the velocity gauge. 
Aversa and Sipe\cite{aversaPRB95} 
presented divergence free expressions for the susceptibility tensor using 
the length gauge that are significantly easier to derive and calculate. 
Cabellos \emph{et al}.\cite{cabellosPRB09} 
elaborated a detailed derivation of the length gauge 
formalism and a comparison with the velocity gauge for bulk GaAs. They 
proved that the results are indeed gauge-invariant. 
This presents a strong case for using the length over the velocity gauge 
for more generalized, and less expensive calculations of the nonlinear
response of a system.

The inclusion of nonlocal effects, such as the scissors correction, has been
correctly included in both the velocity and length guage schemes. 
In the former, both 
Rashkeev \emph{et al}.\cite{rashkeevPRB98} and 
Hughes and Sipe\cite{hughesPRB96} 
accounted for the scissors correction when calculating the nonlinear 
susceptibility tensor.
Leitsmann \emph{et al}.\cite{leitsmannPRB05} 
included excitonic and local field interactions working with GaAs.
Nastos \emph{et al}.\cite{nastosPRB05} demonstrated the correct 
way to calculate the nonlinear response using the scissored Hamiltonian 
in the length gauge, providing better agreement with experiment.

@@

The most complete approach will be a combination of the different techniques 
that lead to more efficient and accurate calculations. Therefore,
the goal of this paper is to present newly derived expressions for the 
nonlinear susceptibility tensor $\chi^{abc}(-2\omega;\omega,\omega)$ 
in the length gauge formalism, 
that include nonlocal terms from the scissors approximation 
and the spatial contribution from the pseudopotentials, and a 
layer-by-layer approach convenient for surface SHG calculations. We are
able to divide the resulting tensor in interband, intraband, $1\omega$, and 
$2\omega$ contributions. This approach is unprecedented and yet produces 
results that are in total agreement with the literature.

The paper is organized as follows. 
In Sec. \ref{theory} we present the 
relevant steps for deriving the surface second-order susceptibility tensor 
within the length gauge formalism. This derivation includes the addition of 
the nonlocal terms mentioned above and a layer-by-layer analysis that have 
been absent in previous works. 
In Sec. \ref{results} we show the results of this 
reformulation with a layer-by-layer study of a clean Si(111) surface with a 
$2 \times 1$ reconstruction. These calculations are done with \emph{ab initio}
programs based on density-functional theory (DFT) within the local-density 
approximation (LDA), using Troullier-Martins pseudopotentials. We compare 
results from before and after adding the different nonlocal contributions.
Finally, in Sec. \ref{conclusions} we submit our conclusions.

>>>>

Numerous mathematical frameworks have been developed to study the 
nonlinear optical response of materials; 
in the context of this paper, we note two in particular. 
The first, called the ``velocity gauge'' or 
``$\mathbf{v}\cdot\mathbf{A}$ gauge,'' features in early 
research in this field. 
Butcher and McLean\cite{butcherPOPS63}, and 
Aspnes\cite{aspnesPRB72} 
laid the initial groundwork by deriving explicit expressions for the 
nonlinear susceptibility tensor. However, these either diverged at 
low frequencies or were limited to zincblende crystal structures. 
A few decades later, 
Ghahramani \emph{et al}.\cite{ghahramaniPRB91} 
realized a full-band-structure calculation for Si/Ge superlattices and 
found significant surface SHG while providing proof that the expressions 
do not diverge at low frequencies for clean, cold semiconductors. 
Sipe \emph{et al}.\cite{sipePRB93}
separated the interband and intraband one-electron contributions 
of the susceptibility tensor and further verified the non-divergent behavior. 
Levine\cite{levinePRB94} 
introduced the scissors operator within the Local Density Approximation
(LDA) and found good agreement with experimental results. 
The second, called the ``length gauge'' or 
``$\mathbf{r}\cdot\mathbf{E}$ gauge'' has various advantages over 
the velocity gauge. 
Aversa and Sipe\cite{aversaPRB95} 
presented divergence free expressions for the susceptibility tensor using 
the length gauge that are significantly easier to derive and calculate. 
Cabellos \emph{et al}.\cite{cabellosPRB09} 
elaborated a detailed derivation of the length gauge 
formalism and a comparison with the velocity gauge for bulk GaAs. They 
proved that the results are indeed gauge-invariant. 
This presents a strong case for using the length over the velocity gauge 
for more generalized and less expensive calculations of the nonlinear
response of a system.

The inclusion of nonlocal effects, such as the scissors correction, has been
correctly included in both the velocity and length guage schemes. 
In the former, both 
Rashkeev \emph{et al}.\cite{rashkeevPRB98} and 
Hughes and Sipe\cite{hughesPRB96} 
accounted for the scissors correction when calculating the nonlinear 
susceptibility tensor.
Leitsmann \emph{et al}.\cite{leitsmannPRB05} 
included excitonic and local field interactions working with GaAs.
Nastos \emph{et al}.\cite{nastosPRB05} demonstrated the correct 
way to calculate the nonlinear response using the scissored Hamiltonian 
in the length gauge providing better agreement with experiment.

@@

The most complete approach will be a combination of the different techniques 
that lead to more efficient and accurate calculations. Therefore,
the goal of this paper is to present newly derived expressions for the 
nonlinear susceptibility tensor $\chi^{abc}(-2\omega;\omega,\omega)$ 
in the length gauge formalism
that include nonlocal terms from the scissors approximation 
and the spatial contribution from the pseudopotentials and a 
layer-by-layer approach convenient for surface SHG calculations. We are
able to divide the resulting tensor in interband, intraband, $1\omega$, and 
$2\omega$ contributions. This approach is unprecedented and yet produces 
results that are in complete agreement with the literature.

The paper is organized as follows. 
In Sec. \ref{theory} we present the 
relevant steps for deriving the surface second-order susceptibility tensor 
within the length gauge formalism. This derivation includes the addition of 
the nonlocal terms mentioned above and a layer-by-layer analysis that have 
been absent in previous works. 
In Sec. \ref{results} we show the results of this 
reformulation with a layer-by-layer study of a clean Si(111) surface with a 
$2 \times 1$ reconstruction. These calculations are done with \emph{ab initio}
programs based on density-functional theory (DFT) within the local-density 
approximation (LDA) using Troullier-Martins pseudopotentials. We compare 
results from before and after adding the different nonlocal contributions.
Finally, in Sec. \ref{conclusions} we submit our conclusions.
